What factors and how influence Zillow Housing Sale Price

### 2024-04-17 paper/paper.qmd
Add files via upload

---
title: "STA302 Project"
subtitle: "What factors and how influence Ames Iowa Housing Sale Price"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: https://github.com/Yuki010305/Final-Paper."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: html
number-sections: true
bibliography: references.bib
---

@@

@@

The housing issue is a pillar industry of a country, and the healthy development of the housing issue affects the country's economic development level. In order to help the stable development of the housing transaction market, this project plans to better evaluate the value of housing transactions for Ames. So, the main purpose of this project is to explore the factors that influence the sale price of homes in Ames, Iowa, and how they affect the home sales price. The characteristics of the homes we selected include central air condition, car capacity of garage, general shape of property, rate of overall quality, total above ground living area, number of fireplaces and sale year. The remainder of this paper is structured as follows. @sec-data....

@@

```{r}
#| label: ames
#| fig-cap: ames house
#| echo: false
#current_dir <- dirname(rstudioapi::getActiveDocumentContext()$path)
#setwd(current_dir)

ames_raw <- read.csv("../data/AmesHousing.csv")

@@

The Raw data collects 82 features of the house, and data quality is not high.

```{r}
#| label: ames select variable
#| fig-cap: ames house save
#| echo: false
ames <- dplyr::select(ames_raw,'Central.Air', 'Lot.Shape', 'Garage.Cars', 'Overall.Qual', 'TotRms.AbvGrd', 'Gr.Liv.Area', 'Garage.Area', 'Fireplaces','Yr.Sold', 'SalePrice')
ames <- na.omit(ames)

write.csv(ames,"../data/analysis_data/ames_analysis.csv",row.names = FALSE)

@@

```{r}
#| label: summary
#| fig-cap: summary numeric
#| echo: false
summary(ames[,-c(1,2)])
```

```{r}
#| label: air
#| fig-cap: central air table
#| echo: false
### Categorical table
table(ames$`Central.Air`)
prop.table(table(ames$`Central.Air`))
```

```{r}
#| label: lot
#| fig-cap: lot shape table
#| echo: false
table(ames$`Lot.Shape`)
prop.table(table(ames$`Lot.Shape`))
```

```{r}
#| label: histprice
#| fig-cap: sale price hist
#| echo: false
par(mfrow=c(1,2))
hist(ames$SalePrice, breaks=10, xlab="Sale Price",main="Sale price",col="red") 
hist(log(ames$SalePrice), breaks=10, xlab="log-transform Sale Price",main="Sale price",col="red") 
```

@@

par(mfrow=c(3,2))

hist(ames$Garage.Area, breaks=10,main="", xlab="Garage Area")
hist(ames$Gr.Liv.Area, breaks=10, main="",xlab="Above grade living area")

boxplot(ames$Overall.Qual,xlab="Overall Qual")
boxplot(ames$TotRms.AbvGrd, xlab="TotRms AbvGrd")

boxplot(ames$Fireplaces, xlab="Fireplaces")
```

```{r,warning=FALSE}
#| label: point
#| fig-cap: points
#| echo: false
a=ggplot(data=ames, aes(x=Garage.Area, y=SalePrice)) + 
  geom_point() + 
  geom_smooth(method = lm, se = FALSE,size=1.5,color="red") + 
  labs(x = 'Garage Area', y='Sale Price')+
  theme_bw()

b=ggplot(data=ames, aes(x=Gr.Liv.Area, y=SalePrice)) + 
  geom_point() + 
  geom_smooth(method = lm, se = FALSE,size=1.5) + 
  labs(x = 'Above grade living area square feet', y='Sale Price')+
  theme_bw()

@@

The above scatter plot shows that both Garage Area and Above grade living area have a certain positive impact on sale price, especially the positive correlation between Above grade living area and sale price is very strong.

@@

```{r}
#| label: spilt
#| fig-cap: spilt data
#| echo: false
#| 
set.seed(123456)
rows <- sample(1:round(0.75*nrow(ames)), replace=FALSE) 
train<- ames[rows,]
test=ames[-rows,]
```

@@

```{r}
#| label: full
#| fig-cap: full model
#| echo: false
#| 
### full model
full = lm(log(SalePrice) ~ ., data = train)
summary(full)
```

```{r}
#| label: sub
#| fig-cap: sub model
#| echo: false
#| 
ml = lm(log(SalePrice) ~ .-TotRms.AbvGrd-Yr.Sold, data = train)
summary(ml)
```

```{r}
#| label: ftest
#| fig-cap: F test
#| echo: false

# F test
anova(ml,full)
```

Analyzing the model of full and significant variables through the Partial F test, we found that the p value is much greater than 0, which shows that we cannot reject the null hypothesis and there is not much difference in the performance of the two models. But in terms of the number of variables, we still choose the model with significant variables.

```{r}
#| label: plot1
#| fig-cap: plot model
#| echo: false
par(mfrow=c(2,2))

plot(ml)
```

@@

# determine whether there are leverage points
n <- nrow(train)
p <- length(coef(ml))-1
# leverage cutoff
h_cut <- 2*(p+1)/n 
off = which(hatvalues(ml) > h_cut)
new_train = train[-off,]
```

@@

```{r}
#| label: model2
#| fig-cap: create model 2
#| echo: false
#| 
ml_new = lm(log(SalePrice) ~ .-TotRms.AbvGrd-Yr.Sold, data =new_train)
summary(ml_new)
```

```{r}
#| label: savemodel
#| fig-cap: save model
#| echo: false
#| 
saveRDS(ml_new, "../model/model.rds")
```

@@

```{r}
#| label: plotmodel2
#| fig-cap: plot model 2
#| echo: false
par(mfrow=c(2,2))
plot(ml_new)
```

@@

```{r}
#| label: plothat
#| fig-cap: plot y hat
#| echo: false
#| 
plot(new_train$SalePrice ~ exp(fitted(ml_new)), main="Y versus Y-hat", xlab="Y-hat", ylab="Y")
abline(a = 0, b = 1)
lines(lowess(new_train$SalePrice ~ exp(fitted(ml_new))), lty=2)
```

@@

@@

homoskedasticity

@@

@@

```{r}
#| echo: false
#| label: read
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

model <-
  readRDS(file = ("../model/model.rds"))

summary(model)
#stargazer(model, type="latex",
#          dep.var.labels=c("Log Sale Price"), align = TRUE)
```

1. When other factors remain unchanged, having central air conditioning increases the sales price of a house by 16.22%.

@@

## First discussion point {#sec-first-point}

## Second discussion point

## Third discussion point

@@

Weaknesses and next steps should also be included.

@@

# Appendix {.unnumbered}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

## Diagnostics

>>>>

---
title: "What factors and how influence Zillow Housing Sale Price"
author: 
  - "Yaning Jin"
thanks: "Code and data are available at: https://github.com/Yuki010305/Final-Paper."
date: today
date-format: long
abstract: "This project is interested in the factors affecting housing prices on zillow. After exploratory analysis of the data, a multiple linear regression was constructed, and a log transformation was performed to build a model with better performance. The transformed housing factors can explain more than 80% of the changes in housing prices. However, the existence of residuals in the model does not fully meet the constraints of the normality assumption, so in the future, data processing, feature engineering and model improvement are needed to study the factors affecting Zillow's housing prices.."
format: pdf
number-sections: true
bibliography: references.bib
link-citations: true
---

@@

\tableofcontents

@@

One of a nation's key industries is housing, and the health of this sector influences the degree of economic growth of the nation. A person's home is frequently the most expensive item in their life. Zillow is an online real estate database provider that assesses property values and offers details on the houses you're interested in [@loukissas2018all]. The Zestimate prediction system was developed by Zillow. It is based on a data collection of hundreds of millions of homes and provides a preliminary estimate of a property's worth by combining a particular algorithm with the features of each home and the state of the market[@rolli2020zillow]. To improve the accuracy of the Zestimate system and provide people with a more trustworthy way to estimate home prices, Zillow launched the Zillow Awards competition in 2017. Our goal in this project is not to participate in a competition to minimize the logarithmic error between the estimated house price and the actual selling price. The main purpose of our exploration in this project is to explore the factors that affect Zillow home sales prices and how they affect home sales prices. Characteristics of the homes we select include number of bathrooms, number of bedrooms, square footage, number of rooms, year, tax value, land tax value, etc.

@@

```{r}
#| label: Zillow
#| fig-cap: Zillow house
#| echo: false

zillow <- read.csv("../data/analysis_data/zillow_price.csv")

@@

The Raw data collects 60 features of the house, and data quality is not high, there are many features with many null values, so 8 high-quality variables are selected from the variables for modeling and prediction.

```{r}
#| label: Zillow select variable
#| fig-cap: Zillow house save
#| echo: false

numeric_cols <- c('bathroomcnt',
                  'bedroomcnt',
                  'calculatedfinishedsquarefeet',
                  'roomcnt',
                  'yearbuilt',
                  'taxvaluedollarcnt',
                  'landtaxvaluedollarcnt',
                  'price')

zillowtrim <- zillow[, numeric_cols]

zillowtrim <- na.omit(zillowtrim)

write.csv(zillowtrim,"../data/analysis_data/zillowtrim_analysis.csv",row.names = FALSE)

@@

|Variable|Description|
|---|---|  
|bathroomcnt|Number of bathrooms in home|   
|bedroomcnt|Number of bedrooms in home|  
|calculatedfinishedsquarefeet|Total finished living area of the home|  
|roomcnt|Total number of rooms|  
|yearbuilt|The Year the principal residence was built|  
|taxvaluedollarcnt|The total tax assessed value of the parcel|  
|landtaxvaluedollarcnt|The assessed value of the land area| 
|price|price|

```{r}
#| label: summary
#| fig-cap: summary numeric
#| echo: false
summary(zillowtrim[,-c(1,2,4,5)])
```

```{r,echo=FALSE,eval=FALSE}
#| label: air
#| fig-cap: central air table
#| echo: false
### Categorical table
table(zillowtrim$bathroomcnt)
```

```{r,echo=FALSE,eval=FALSE}
#| label: bedroomcnt
#| fig-cap: bedroomcnt table
#| echo: false
table(zillowtrim$bedroomcnt)
prop.table(table(zillowtrim$bedroomcnt))
```

```{r}
#| label: histprice
#| fig-cap: sale price hist
#| echo: false
par(mfrow=c(1,2))
hist(zillowtrim$price, breaks=10, xlab="Sale Price",main="Sale price",col="red") 
hist(log(zillowtrim$price), breaks=10, xlab="log-transform Sale Price",main="Sale price",col="red") 
```

@@

par(mfrow=c(1,2))

hist(zillowtrim$bathroomcnt, breaks=10,main="", xlab="bathroom")
hist(zillowtrim$bedroomcnt, breaks=10, main="",xlab="bedroom")

par(mfrow=c(1,2))
boxplot(zillowtrim$calculatedfinishedsquarefeet,xlab="square feet")
boxplot(zillowtrim$roomcnt, xlab="room")

par(mfrow=c(1,2))
boxplot(zillowtrim$taxvaluedollarcnt, xlab="tax value")
boxplot(zillowtrim$landtaxvaluedollarcnt, xlab="land tax value")
```

```{r,warning=FALSE,message=FALSE}
#| label: point
#| fig-cap: points
#| echo: false
a=ggplot(data=zillowtrim, aes(x=bathroomcnt, y=price)) + 
  geom_point() + 
  geom_smooth(method = lm, se = FALSE,size=1.5,color="red") + 
  labs(x = 'bathroom', y='Sale Price')+
  theme_bw()

b=ggplot(data=zillowtrim, aes(x=taxvaluedollarcnt, y=price)) + 
  geom_point() + 
  geom_smooth(method = lm, se = FALSE,size=1.5) + 
  labs(x = 'tax value', y='Sale Price')+
  theme_bw()

@@

The above scatter plot shows that tax value has a certain positive impact on sale price, the positive correlation between tax value and sale price is very strong.

```{r,warning=FALSE,message=FALSE}
#| label: point1
#| fig-cap: points1
#| echo: false
a=ggplot(data=zillowtrim, aes(x=calculatedfinishedsquarefeet, y=price)) + 
  geom_point() + 
  geom_smooth(method = lm, se = FALSE,size=1.5,color="red") + 
  labs(x = 'square feet', y='Sale Price')+
  theme_bw()

b=ggplot(data=zillowtrim, aes(x=bedroomcnt, y=price)) + 
  geom_point() + 
  geom_smooth(method = lm, se = FALSE,size=1.5) + 
  labs(x = 'bedroom', y='Sale Price')+
  theme_bw()

The above scatter plot shows that square feet has a certain positive impact on sale price, the positive correlation between square feet and sale price is very strong.

@@

```{r}
#| label: spilt
#| fig-cap: spilt data
#| echo: false
#| 
set.seed(123456)
rows <- sample(1:round(0.75*nrow(zillowtrim)), replace=FALSE) 
train<- zillowtrim[rows,]
test=zillowtrim[-rows,]
```

@@

```{r}
#| label: full
#| fig-cap: full model
#| echo: false
#| 
### full model
full = lm(log(price) ~ ., data = train)
summary(full)
```

```{r}
#| label: sub
#| fig-cap: sub model
#| echo: false
#| 
log_ml = lm(log(price) ~ bathroomcnt+bedroomcnt+log(calculatedfinishedsquarefeet)+roomcnt+yearbuilt+log(taxvaluedollarcnt)+log(landtaxvaluedollarcnt), data = train)
summary(log_ml)
```

```{r}
#| label: savebestmodel
#| fig-cap: save besr model
#| echo: false
#| 
saveRDS(log_ml , "../models/best_model.rds")
```

```{r}
#| label: plot1
#| fig-cap: plot model
#| echo: false
#par(mfrow=c(2,2))

#plot(log_ml)
```

@@

# determine whether there are leverage points
n <- nrow(train)
p <- length(coef(log_ml))-1
# leverage cutoff
h_cut <- 2*(p+1)/n 
off = which(hatvalues(log_ml) > h_cut)
new_train = train[-off,]
```

@@

```{r}
#| label: model2
#| fig-cap: create model 2
#| echo: false
#| 
ml_new = lm(log(price) ~ bathroomcnt+bedroomcnt+log(calculatedfinishedsquarefeet)+roomcnt+yearbuilt+log(taxvaluedollarcnt)+log(landtaxvaluedollarcnt), data = new_train)
summary(ml_new)
```

```{r}
#| label: savemodel
#| fig-cap: save model
#| echo: false
#| 
saveRDS(ml_new, "../models/model.rds")
```

@@

```{r}
#| label: plotmodel2
#| fig-cap: plot model 2
#| echo: false
plot(ml_new)
```

@@

```{r}
#| label: plothat
#| fig-cap: plot y hat
#| echo: false
#| 
plot(new_train$price ~ exp(fitted(ml_new)), main="Y versus Y-hat", xlab="Y-hat", ylab="Y")
abline(a = 0, b = 1)
lines(lowess(new_train$price ~ exp(fitted(ml_new))), lty=2)
```

@@

the errors are independent.

@@

the errors have constant variance.

@@

```{r}
#| label: plotqq
#| fig-cap: plot qq
#| echo: false
#| 
qqnorm(resid(ml_new))
qqline(resid(ml_new))
```

@@

Estimating the value of a home is a common difficulty. Therefore, a lot of work has already been completed. Lee [@lee2016software] make an effort to develop multivariate regression models based on house datasets and assess the models using maximum information coefficient statistics based on anticipated values and home prices.  When a moderate to big data sample size is employed, Nghiep et al. [@nguyen2001predicting] examined two methods: multiple regression analysis (MRA) and artificial neural networks (ANN). Based on the prediction performance, ANN performs better than MRA. An artificial neural network (ANN) model was created after Limsombunchai [@limsombunchao2004house
] that ANNs are more accurate in predicting property values than hedonic regression models.

The final model shows that the number of bathrooms, bedrooms, rooms, year taxes and fees of the building, square feet, etc. all have a significant impact on Zillow’s housing prices.For every additional bathroom, the house price increases by 1.017% per unit. For every additional bedroom, the house price actually decreases by 0.01%. In addition, the newer the building is, the housing prices actually decrease. For every 1% unit increase in taxes and fees, housing prices increase by 0.67% unit.These findings explain the relationship between house-related attributes and house prices.

@@

@@

However, the residuals of our model do not fully satisfy the normality assumption. Such violations may reduce the model's predictive accuracy on new data sets. Another limitation of this model is that transforming the data makes the model less interpretable.

To start making these models better in the future, divide each training dataset into smaller training tests and perform some cross-validation before generating the test files. This might enhance performance, or at the very least increase the predictability of the model's output. Although most categorical variables cannot be included in the model because of memory and time restrictions when running the model, the model may be enhanced with improved feature engineering. These variables can be included in the model by being divided into smaller groups. Making some interactive words is an additional thought. It is crucial to classify missing variables as predictors once all missing values have been imputed, as was previously noted.

@@
